# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data
```{r ex1,echo=TRUE}
library(lattice)
library(data.table)
library(plyr)

data <- read.csv("activity.csv",header=TRUE,sep=",",stringsAsFactors=FALSE)

data$formattedDate <- as.Date(data[,2],format="%Y-%m-%d")
data$weekDay <- weekdays(data[,4])
```

## What is mean total number of steps taken per day?
```{r ex2,echo=TRUE}
hist(data$steps)

sumSteps <- tapply(data$steps,data$formattedDate,sum)

hist(sumSteps)

auxData <- data
auxData <- auxData[!is.na(auxData$steps),]
```

### Values of the "mean" total number of steps taken per day:
```{r ex2.1,echo=TRUE}
meanSteps <- tapply(auxData$steps,auxData$formattedDate,mean)
meanSteps
#xtMean <- xtable(meanSteps)
#print(xtMean, type="html")
```

### Values of the "median" total number of steps taken per day:
```{r ex2.2,echo=TRUE}
medianSteps <- tapply(auxData$steps,auxData$formattedDate,median)
medianSteps
#xtMedian <- xtable(medianSteps)
#print(xtMedian, type="html")
```


## What is the average daily activity pattern?
```{r ex3,echo=TRUE}

weekendIndex <- data$weekDay == "sábado" | data$weekDay == "domingo"
weekdayIndex <- data$weekDay != "sábado" & data$weekDay != "domingo"
data$weekDayOrWeekEnd[weekendIndex] <- "weekend"
data$weekDayOrWeekEndId[weekendIndex] <- 2
data$weekDayOrWeekEnd[weekdayIndex] <- "weekday"
data$weekDayOrWeekEndId[weekdayIndex] <- 1

data_DT <- data.table(auxData)
data_DT[,date:=NULL]
data_DT[,formattedDate:=NULL]
data_DT[,weekDay:=NULL]
data_DT[,weekDayOrWeekEnd:=NULL]

intervalsMean <- data_DT[,lapply(.SD,mean),by='interval']

plot(x=intervalsMean$interval,y=intervalsMean$steps,type="l",xlab="5-minute Interval",ylab="Avg Number of steps across all days",main="Time series: Avg nº of steps taken\naveraged across all days")
```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r ex3.1,echo=TRUE}
maxSteps <- max(intervalsMean$steps)
maxIndex <- which(intervalsMean$steps == maxSteps)
```

The 5-minute interval with the maximum number of steps:
```{r ex3.2,echo=TRUE}
intervalsMean$interval[maxIndex]
```


## Imputing missing values
### Calculate and report the total number of missing values in the dataset
```{r ex4,echo=TRUE}
dataClean <- data
stepsNA <- is.na(dataClean$steps)
rowsNA <- stepsNA[stepsNA == TRUE]
```

The number of missing values in the dataset:
```{r ex4.1,echo=TRUE}
length(rowsNA)
```

### Fill in the missing values and create a new dataset (using previously calculated 5-minute interval means):
```{r ex4.2,echo=TRUE}
dataClean <- data
replaceData <- intervalsMean
replaceData$steps5minAvg <- replaceData$steps
replaceData[,weekDayOrWeekEndId:=NULL]
replaceData[,steps:=NULL]
dfList <- list(dataClean,replaceData)
dataClean <- join_all(dfList)
dataClean$steps[stepsNA] <- dataClean$steps5minAvg
```

### Make histograms (after imputing missing values):
```{r ex4.3,echo=TRUE}
hist(dataClean$steps)

sumSteps <- tapply(dataClean$steps,dataClean$formattedDate,sum)

hist(sumSteps)
```

### Values of the "mean" total number of steps taken per day (after imputing missing values):
```{r ex4.4,echo=TRUE}
meanSteps <- tapply(dataClean$steps,dataClean$formattedDate,mean)
meanSteps
```

### Values of the "median" total number of steps taken per day (after imputing missing values):
```{r ex4.5,echo=TRUE}
medianSteps <- tapply(dataClean$steps,dataClean$formattedDate,median)
medianSteps
```


## Are there differences in activity patterns between weekdays and weekends?
```{r ex5,echo=TRUE}
dataClean_DT <- data.table(dataClean)
dataClean_DT[,date:=NULL]
dataClean_DT[,formattedDate:=NULL]
dataClean_DT[,weekDay:=NULL]
dataClean_DT[,weekDayOrWeekEnd:=NULL]

intervalsMean <- dataClean_DT[,lapply(.SD,mean),by='interval,weekDayOrWeekEndId']

weekDayIndex <- intervalsMean$weekDayOrWeekEndId == 1
weekEndIndex <- intervalsMean$weekDayOrWeekEndId == 2
intervalsMean$weekDay[weekDayIndex] <- "WeekDay"
intervalsMean$weekDay[weekEndIndex] <- "WeekEnd"

xyplot(intervalsMean$steps ~ intervalsMean$interval | intervalsMean$weekDay, layout = c(1, 2),type="l",xlab="Interval",ylab="Number of steps")
```